An algorithm for a spatial mapping using a hexagon tilegram, with application to Australia disease maps

### 2019-02-21 algorithm_radix.Rmd
included buffer code

```{r libraries}
library(tidyverse)
# To install:
# devtools::install_github("srkobakian/sugaRbag")
library(sugaRbag)
```

The polygon set of Local Government Areas in Australia in 2011 is stored in the sugaRbag package as `lga_2011`.

```{r data}
sf_id <- "LGA_CODE11"
focal_points <- sugaRbag::capital_cities
```

@@

```{r, eval = FALSE}
# Read in shape information
shp <- read_shape(shp_path = shp_path)

# Convert to a simple features object
shp_sf <- sf::st_as_sf(shp)
```

@@

We will use the Local Government Areas of Australia in the `lga_2011` data set, and take only those from New South Wales as an example:

```{r}
shp_sf <- sugaRbag::lga_2011

nsw_lga <- shp_sf %>% dplyr::filter(STE_NAME11 == "New South Wales")
# Select only desired LGA areas, those from nsw.
nsw_lga[[sf_id]] <- droplevels(as.factor(nsw_lga[[sf_id]]))
```

```{r project}
## Not necessary when using `create_hexmap`:
## Derive projection information
crs_info <- sf::st_crs(nsw_lga)
proj4string <- crs_info$proj4string
epsg <- crs_info$epsg

## For UTM as per shape file
projstring <- paste0("+init=epsg:", epsg, " ", proj4string, collapse = "")
## if long lat is preferred to UTM, and not in proj4string:
projstring <- "+init=epsg:3112 +proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=0 +lon_0=134 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"

nsw_lga <- sf::st_transform(nsw_lga, crs = projstring)
```

```{r area_variable}
nsw_lga <- nsw_lga %>% mutate(logarea = log(AREA_SQKM))
```

```{r centroids}
## Use the create_centroids function to find polygon centroids:
centroids <- create_centroids(shp_sf = nsw_lga, sf_id = sf_id, projstring = projstring, verbose = FALSE)

centroids$logarea <- nsw_lga$logarea

fort_nsw <- fortify_sfc(nsw_lga, projstring = projstring)
```

```{r, fig.cap="NSW Local Government Areas and derived centroids."}
fort_nsw <- fortify_sfc(nsw_lga, projstring = projstring)

end_cents <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude, colour = logarea), data= centroids) + theme_void() + coord_equal()
end_cents
```

@@

```{r plot_grid}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r plot_shift}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r grid_id}
full_grid <- grid %>%
    mutate(hex_long_int = dense_rank(hex_long)-1,
        hex_lat_int = dense_rank(hex_lat)-1)
```

@@

To filter the grid for appropriate hexagons for allocation,
the `create_buffer` function is used.

```{r buffer}
grid <- create_buffer(centroids, full_grid, hex_size, buffer_dist)
```

@@

The first rolling window function finds the minimum and maximum centroid values for the groups of longitude columns and the groups of latitude rows.

@@

@@

```{r plot_buffer}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a", fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = full_grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75)
```

@@

```{r, eval = FALSE}
hexmap_allocation <- allocate(centroids = centroids,
        hex_grid = grid,
        hex_size = hex_size,
        filter_dist = filter_dist,
        width = 15,
        focal_points = focal_points,
        verbose = TRUE,
        id = sf_id)
```

@@

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

@@

ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

```{r}
hex_points_df <- centroid_allocation %>% 
  group_by(LGA_CODE11) %>%
  nest() %>%
  mutate(hex_points = 
      map(.x = data,.f = function(a) { 
        fortify_hexagon(data = a, hex_size = hex_size)})) %>%
  unnest(data) %>%
  unnest(hex_points)
```

```{r}
ggplot(hex_points_df) + 
  geom_polygon(aes(x = hexv_long, y = hexv_lat, group = LGA_CODE11)) +
  coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

```{r fullhexmap}
lga_map <- create_hexmap(shp = nsw_lga, sf_id = "LGA_CODE11", buffer_dist = NULL, filter_dist = 10, export_shp = FALSE, focal_points = sugaRbag::capital_cities, projstring = projstring, verbose = TRUE)

hexmap_lga <- left_join(nsw_lga, 
  lga_map)

@@

```{r}
end_hex <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_nsw, fill = "grey", colour = "white") +
  theme_void() + coord_equal() +
    geom_hex(aes(x = hex_long, y = hex_lat, fill = AREA_SQKM), data = hexmap_lga, 
      position = "identity", stat = "identity")

@@

Examples of Melbourne, nsw, Australia
At different levels, distance from original centroids, change in area

LGA of all Australia
SA2 for nsw

>>>>

```{r libraries}
library(tidyverse)
# To install sugaRbag:
# devtools::install_github("srkobakian/sugaRbag")
library(sugaRbag)
```

The polygon set of Local Government Areas in Australia in 2011 is stored in the sugaRbag package data as `tas_sa2`. To create this object:

```
# Download resource 
# link to folder containing .shp file on sugaRbag site.

aus_sa2 <- read_shape("file location")

filter for Tasmanian areas

tas_sa2 <- filter(aus_sa2, aus_sa2$STE_NAME11 == "Tasmania)
```

A column of the data set must be used as a unique identifier of areas.

```{r data}
sf_id <- "SA2_5DIG16"
focal_points <- sugaRbag::capital_cities
```

@@

<!--
```
# Read in shape information
shp <- read_shape(shp_path = shp_path)
```
-->

@@

We will use the Statical Areas (Level 2) of Australia, provided by Australian Bureau of Statistics. With only those from Tasmania as an example:

```{r}
# Convert to a simple features object
tas_sa2 <- sf::st_as_sf(tas_sa2)

# Select only desired LGA areas, those from tas.
tas_sa2[[sf_id]] <- droplevels(as.factor(tas_sa2[[sf_id]]))

fort_tas <- fortify_sfc(tas_sa2)
```

```{r}
# Include area variable
area_var <- fort_tas %>% 
  distinct(SA2_5DIG16, .keep_all = TRUE) %>% 
  select(SA2_5DIG16, AREASQKM16)
```

```{r centroids}
## Use the create_centroids function to find polygon centroids:
centroids <- create_centroids(shp_sf = tas_sa2, sf_id = sf_id) %>%
  left_join(., select(area_var, SA2_5DIG16, AREASQKM16), by = "SA2_5DIG16")
```

```{r, fig.cap="tas Local Government Areas and derived centroids."}
end_cents <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude, colour = log(AREASQKM16)), data= centroids) + theme_void() + coord_equal()
end_cents

#ggplot() + geom_histogram(aes(x = log(AREASQKM16), fill = log(AREASQKM16)), data= centroids, bins = 20) 
```

@@

```{r plot_grid}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude), data= centroids, colour = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r plot_shift}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75) 
```

@@

```{r grid_id}
full_grid <- grid <- grid %>%
    mutate(hex_long_int = dense_rank(hex_long)-1,
        hex_lat_int = dense_rank(hex_lat)-1)
```

@@

To filter the grid for appropriate hexagons for allocation,
the `create_buffer` function is called from `create_grid`.

It finds the amount of columns and rows needed to capture the set of centroids.

```{r}
nlong <- length(unique(grid$hex_long))
nlat <- length(unique(grid$hex_lat))

centroids <- centroids %>%
  mutate(long_int = round((longitude-min(grid$hex_long))/(max(grid$hex_long)-min(grid$hex_long))*nlong, 0),
          lat_int = round((latitude-min(grid$hex_lat))/(max(grid$hex_lat)-min(grid$hex_lat))*nlat, 0))
```

@@

```{r echo = FALSE}
# Amount of lats and longs in each group
lat_size = round(nlat/20,0)
long_size = round(nlong/20,0)
```

The first rolling window function finds the minimum and maximum centroid values for the groups of longitude columns and the groups of latitude rows.
The specific rows and columns in the rolling windows are defined.

```{r echo = FALSE}
 # make a list of groups, manual sliding windows
    nlat_list <- purrr::map2(seq(1:nlat), lat_size + seq(1:nlat), c)
    nlong_list <- purrr::map2(seq(1:nlong), long_size + seq(1:nlong), c)

    lat_window <- function(x, cents = centroids, maximum = nlat){
        max_int = min(x[2],maximum)

        cents_in <- filter(cents, between(lat_int, x[1], max_int))
        return(cents_in)
    }

    long_window <- function(x, cents = centroids, maximum = nlong){
        max_int = x[2]
        while (max_int > maximum){
            max_int = max_int - 1
        }

        cents_in <- filter(cents, between(long_int, x[1], max_int))
        return(cents_in)
    }

    # LATITUDE ROWS FILTER
    # amount of latitude in sliding window
    lat_windows <- purrr::map(.x = nlat_list, .f = lat_window)

    # find the min and max longitude for each latitude
    range_rows <- purrr::map_dfr(.x = lat_windows,
        .f = function(x) {x %>%
                dplyr::summarise(
                    long_min = ifelse(purrr::is_empty(long_int), NA, min(x$long_int)),
                    long_max = ifelse(purrr::is_empty(long_int), NA, max(x$long_int))
                )}
    )
```

@@

```{r echo = FALSE}
 # smooth the minimums
    av_range_rows <- purrr::map_dfr(.x = nlat_list, .f = function(x, rows = range_rows) {
        rows[x[1]:min(x[2], NROW(rows)),] %>%
            dplyr::summarise(mean_long_min = mean(long_min, na.rm=T), mean_long_max = mean(long_max, na.rm=T))
    }) %>%
        bind_cols(lat_id = c(seq(1:nlat) +lat_size), .)

    # LONGITUDE COLS FILTER
    long_windows <- purrr::map(.x = nlong_list, .f = long_window, centroids, nlong)

    # find the min and max longitude for each latitude
    range_cols <- purrr::map_dfr(.x = long_windows, .f = function(x) { x %>%
            dplyr::summarise(
                lat_min = ifelse(purrr::is_empty(lat_int), NA, min(x$lat_int)),
                lat_max = ifelse(purrr::is_empty(lat_int), NA, max(x$lat_int))
            )}
    )

    # smooth the minimums
    av_range_cols <- purrr::map_dfr(.x = nlong_list, .f = function(x, cols = range_cols) {
        cols[x[1]:min(x[2], NROW(cols)),] %>%
            dplyr::summarise(mean_lat_min = mean(lat_min, na.rm=T), mean_lat_max = mean(lat_max, na.rm=T))
    }) %>%
        bind_cols(long_id = c(seq(1:nlong) + round(long_size/2)), .)

@@

```{r echo = FALSE}
    # APPLY A BUFFER
    # change buffer to amount of hexagons (ints) either side
    hex_buffer <- floor(buffer_dist/hex_size)

    grid <- grid %>%
        left_join(., av_range_rows, by = c("hex_lat_int" = "lat_id")) %>%
        left_join(., av_range_cols, by = c("hex_long_int" = "long_id")) %>%
        rowwise() %>%
        mutate(long_buffer = ifelse(between(hex_long_int,mean_long_min - hex_buffer,
            mean_long_max + hex_buffer), "in", "out")) %>%
        mutate(lat_buffer = ifelse(between(hex_lat_int,mean_lat_min - hex_buffer,
            mean_lat_max + hex_buffer), "in", "out")) %>%
        filter(lat_buffer =="in" | long_buffer == "in")

```{r plot_buffer}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroids, fill = "#b2df8a", fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = full_grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = grid, size = 0.75)
```

@@

```{r}
width = 15
```

```{r, eval = FALSE}
hexmap_allocation <- allocate(centroids = centroids,
        hex_grid = grid,
        hex_size = hex_size,
        filter_dist = filter_dist,
        width = width,
        focal_points = focal_points,
        verbose = TRUE,
        id = sf_id)
```

@@

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

```{r, echo = FALSE}
ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

@@

ggplot() + geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") + geom_point(aes(x=longitude, y = latitude),data= centroid1, fill = "red") + theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)

@@

```{r}
hex_points_df <- centroid_allocation %>% 
  group_by(SA2_CODE11) %>%
  nest() %>%
  mutate(hex_points = 
      map(.x = data,.f = function(a) { 
        fortify_hexagon(data = a, hex_size = hex_size)})) %>%
  unnest(data) %>%
  unnest(hex_points)
```

```{r}
ggplot(hex_points_df) + 
  geom_polygon(aes(x = hexv_long, y = hexv_lat, group = SA2_CODE11)) +
  coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#a6cee3", data = grid, size = 0.25) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#1f78b4", data = hex_grid, size = 0.75)
```

```{r fullhexmap}
lga_map <- create_hexmap(shp = tas_sa2, sf_id = "SA2_CODE11", buffer_dist = NULL, filter_dist = 10, export_shp = FALSE, focal_points = sugaRbag::capital_cities, projstring = projstring, verbose = TRUE)

hexmap_lga <- left_join(tas_sa2, 
  lga_map)

@@

```{r}
end_hex <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), data = fort_tas, fill = "grey", colour = "white") +
  theme_void() + coord_equal() +
    geom_hex(aes(x = hex_long, y = hex_lat, fill = AREA_SQKM), data = hexmap_lga, 
      position = "identity", stat = "identity")

@@

Examples of Melbourne, tas, Australia
At different levels, distance from original centroids, change in area

LGA of all Australia
SA2 for tas
